Further Empirical Analyses of Rawlsian Fairness for Machine Learning

### 2017-05-09 paper/paper.tex
added conclusion

\title{Title}

@@

\begin{enumerate}
	\item We read the paper \textit{Rawlsian Fairness for Machine Learning}  \cite{DBLP:journals/corr/JosephKMNR16}.
	\item We implemented the \texttt{TopInterval}, \texttt{IntervalChaining}, and \texttt{RidgeFair} algorithms from the paper in Python. 
	\item We ran our implementations on a Yahoo! dataset containing a fraction of the user click log for news articles displayed in the Featured Tab of the Today Module on the Yahoo! Front Page during the first ten days in May 2009 \cite{yahoo}, to see how well they performed on real data.
	\item To empirically evaluate our implementations, we ran experiments similar to those in \cite{DBLP:journals/corr/JosephKMNR16} with randomly-drawn contexts. 
	\item We compiled our findings into a written report.
\end{enumerate}

@@

The \texttt{TopInterval} learning algorithm was implemented true to form as presented in \cite{DBLP:journals/corr/JosephKMNR16}. Particular details of note --- to ensure that all matrices used in computation were nonsingular, the first $d$ rounds are always chosen to be exploration rounds, where $d$ is the number of features. Additionally, we found it necessary to pick each arm once in order to observe data for each.

@@

The implementation for \texttt{IntervalChaining} was simple given \texttt{TopInterval}, as it sufficed to alter the strategy for picking arms in each round to that of picking uniformly at random from the chain containing the top interval.

@@

The \texttt{RidgeFair} algorithm was also implemented as presented in \newcite{DBLP:journals/corr/JosephKMNR16}. This algorithm is very similar in implementation to \texttt{IntervalChaining}, save that it’s narrower confidence intervals allow for derivation of tighter regret bounds. A couple small details to note are, first, that we assume for simplicity (and without loss of generality) that the noise is $R$ sub-Gaussian with parameter $R = 1$ and second, that we play uniformly at random among all arms in the set of actions chained to the max.

@@

Initial attempts to use this approach failed for a couple of reasons. First, the Yahoo! dataset contains a highly disproportionate number of negative samples with respect to positive ones. Therefore, our learning algorithm would not retain useful information over a number of iterations due to only being trained on negative samples. Second, a direct application of the \texttt{TopInterval} and \texttt{IntervalChaining} algorithms relies on the assumption of $20$ distinct underlying groups from which the articles were chosen to be in the article pool, each with their own distinct quality function. This assumption was found to be unreasonable, as we found that an article's index in the article pool had no bearing on its actual likelihood of being clicked by the user when picked. The initial context also does not work well with a fairness analysis. As a consequence, we saw that direct applications of the learning algorithms saw very poor performance.

@@

Another novel modification we made was the use of a logit model instead of the simple linear regression used in \newcite{DBLP:journals/corr/JosephKMNR16}. We preserve the original fairness argument of the \texttt{IntervalChaining} algorithm by simply rescaling the output of the OLS estimator and the confidence intervals to $[0, 1]$ via the inverse logit. That is,
$$w_{t,i} = \mathcal{Q}_{\mathcal{F}_{t,i}}(\frac{\delta}{2kT})$$
$$[\ell_{i}^{t}, u_{i}^{t}] = [\Phi(\hat{y}_{t,i} - w_{t,i}), \Phi(\hat{y}_{t,i} + w_{t,i})]$$
where $\Phi(x) = \frac{e^{x}}{1 + e^{x}} = \text{logistic}(x)$. It suffices to note that both OLS and logistic regression are variations of the generalized linear model (GLM).

@@

We ran experiments that compared the regret of \textsc{IntervalChaining} (IC) with the regret of \textsc{TopInterval} (TI). As in \newcite{DBLP:journals/corr/JosephKMNR16}, we present three sets of empirical results: 
\begin{itemize}
	\item Varying $T$ (the number of rounds) - we measured the average regret of \textsc{IntervalChaining} and \textsc{TopInterval} as a function of increasing $T$. (Figure \ref{fig:free_T})
	\item Varying $k$ (the number of arms/groups) - we measured the average regret of \textsc{IntervalChaining} and \textsc{TopInterval} as a function of increasing $k$. (Figure \ref{fig:free_k})
	\item Varying $d$ (the number of features) - we measured the average regret of \textsc{IntervalChaining} and \textsc{TopInterval} as a function of increasing $d$. (Figure \ref{fig:free_d})
\end{itemize}

@@

>>>>

\title{Further Empirical Analyses of Rawlsian Fairness for Machine Learning}

@@

\begin{enumerate}
	\item We read the paper \textit{Rawlsian Fairness for Machine Learning}  \cite{DBLP:journals/corr/JosephKMNR16}.
	\item We implemented the \textsc{TopInterval}, \textsc{IntervalChaining}, and \textsc{RidgeFair} algorithms from the paper in Python. 
	\item We ran our implementations on a Yahoo! dataset containing a fraction of the user click log for news articles displayed in the Featured Tab of the Today Module on the Yahoo! Front Page during the first ten days in May 2009 \cite{yahoo}, to see how well they performed on real data.
	\item To empirically evaluate our implementations, we ran experiments similar to those in \cite{DBLP:journals/corr/JosephKMNR16} with randomly-drawn contexts. 
	\item We compiled our findings into a written report.
\end{enumerate}

@@

The \textsc{TopInterval} learning algorithm was implemented true to form as presented in \newcite{DBLP:journals/corr/JosephKMNR16}. Particular details of note --- to ensure that all matrices used in computation were nonsingular, the first $d$ rounds are always chosen to be exploration rounds, where $d$ is the number of features. Additionally, we found it necessary to pick each arm once in order to observe data for each.

@@

The implementation for \textsc{IntervalChaining} was simple given \textsc{TopInterval}, as it sufficed to alter the strategy for picking arms in each round to that of picking uniformly at random from the chain containing the top interval.

@@

The \textsc{RidgeFair} algorithm was also implemented as presented in \newcite{DBLP:journals/corr/JosephKMNR16}. This algorithm is very similar in implementation to \textsc{IntervalChaining}, save that it’s narrower confidence intervals allow for derivation of tighter regret bounds. A couple small details to note are, first, that we assume for simplicity (and without loss of generality) that the noise is $R$ sub-Gaussian with parameter $R = 1$ and second, that we play uniformly at random among all arms in the set of actions chained to the max.

@@

Initial attempts to use this approach failed for a couple of reasons. First, the Yahoo! dataset contains a highly disproportionate number of negative samples with respect to positive ones. Therefore, our learning algorithm would not retain useful information over a number of iterations due to only being trained on negative samples. Second, a direct application of the \textsc{TopInterval} and \textsc{IntervalChaining} algorithms relies on the assumption of $20$ distinct underlying groups from which the articles were chosen to be in the article pool, each with their own distinct quality function. This assumption was found to be unreasonable, as we found that an article's index in the article pool had no bearing on its actual likelihood of being clicked by the user when picked. The initial context also does not work well with a fairness analysis. As a consequence, we saw that direct applications of the learning algorithms saw very poor performance.

@@

Another novel modification we made was the use of a logit model instead of the simple linear regression used in \newcite{DBLP:journals/corr/JosephKMNR16}. We preserve the original fairness argument of the \textsc{IntervalChaining} algorithm by simply rescaling the output of the OLS estimator and the confidence intervals to $[0, 1]$ via the inverse logit. That is,
$$w_{t,i} = \mathcal{Q}_{\mathcal{F}_{t,i}}(\frac{\delta}{2kT})$$
$$[\ell_{i}^{t}, u_{i}^{t}] = [\Phi(\hat{y}_{t,i} - w_{t,i}), \Phi(\hat{y}_{t,i} + w_{t,i})]$$
where $\Phi(x) = \frac{e^{x}}{1 + e^{x}} = \text{logistic}(x)$. It suffices to note that both OLS and logistic regression are variations of the generalized linear model (GLM).

@@

We ran experiments that compared the regret of \textsc{IntervalChaining} (IC) with the regret of \textsc{TopInterval} (TI). As in \newcite{DBLP:journals/corr/JosephKMNR16}, we present three sets of empirical results: 
\begin{itemize}
	\item Varying $T$ (the number of rounds) - we measured the average regret of \textsc{IntervalChaining} and \textsc{TopInterval} as a function of increasing $T$. (See Figure \ref{fig:free_T}.)
	\item Varying $k$ (the number of arms/groups) - we measured the average regret of \textsc{IntervalChaining} and \textsc{TopInterval} as a function of increasing $k$. (See Figure \ref{fig:free_k}.)
	\item Varying $d$ (the number of features) - we measured the average regret of \textsc{IntervalChaining} and \textsc{TopInterval} as a function of increasing $d$. (See Figure \ref{fig:free_d}.)
\end{itemize}

@@

In this work, we present an empirical extension of the work done by \newcite{DBLP:journals/corr/JosephKMNR16} in their paper \textit{Rawlsian Fairness for Machine Learning}. Specifically, we present implementations of their algorithms \textsc{TopInterval}, \textsc{IntervalChaining}, and \textsc{RidgeFair}; a case study in which we apply the aforementioned algorithms to a Yahoo! clicks dataset; and an extension of one of \newcite{DBLP:journals/corr/JosephKMNR16}'s empirical analyses on randomly generated data. We believe that both our results may be useful should these algorithms be used in future real-world settings.
